Recurrent Deep Learning Models and its applications

### 2025-02-16 paper2/main.tex
delete conda file

>>>>

To achieve the project goals and objectives, the research will follow a structured approach combining theoretical understanding with empirical evaluation:

1. **Literature Review:**
   - **Background Study:**  
     Conduct a detailed review of seminal and recent works on RNNs, LSTMs, GRUs, and RNN encoder-decoder architectures in NLP. Key references include:
     - Hochreiter \& Schmidhuber (1997) on LSTM.
     - Cho et al. (2014) and Chung et al. (2014) on GRU.
     - Recent surveys on the application of RNNs in NLP.
   - **Problem Context:**  
     Understand current best practices, challenges (e.g., vanishing/exploding gradients, sequence length handling), and evaluation standards in NLP.

2. **Data Collection and Preprocessing:**
   - **Dataset Selection:**  
     Choose a benchmark NLP dataset appropriate to the target task (e.g., language modeling, text classification, or machine translation).
   - **Preprocessing Steps:**  
     Perform text normalization, tokenization, vocabulary building, and sequence padding. Optional steps may include initializing word embeddings (either pre-trained or learned from scratch).

3. **Model Development and Experimental Implementation:**
   - **Model Architecture Design:**  
     Implement the different RNN-based models:
     - **Vanilla RNN:**  
       Define the recurrence as  
       \[
       h_t = \sigma(W_{hx} x_t + W_{hh} h_{t-1} + b_h), \quad y_t = f(W_{yh} h_t + b_y),
       \]
       where \(x_t\) is the input vector at time \(t\), \(h_t\) is the hidden state, and \(y_t\) is the output.
     - **LSTM:**  
       Set up the LSTM cell with input, forget, and output gates to control information flow.
     - **GRU:**  
       Implement the GRU cell, which uses update and reset gates as a simplified alternative.
     - **RNN Encoder-Decoder:**  
       Develop an encoder to summarize the input sequence and a decoder to generate the target sequence.
   - **Framework and Tools:**  
     Develop the models using PyTorch or TensorFlow. Leverage available libraries for reproducibility and efficiency.

4. **Training, Hyperparameter Tuning, and Evaluation:**
   - **Training Protocol:**  
     Use similar training configurations (e.g., optimizer choice, learning rate schedule, batch size) for all models to ensure a fair comparison.
   - **Hyperparameter Optimization:**  
     Conduct grid search or Bayesian optimization for parameters such as hidden unit size, dropout rates, and learning rates.
   - **Evaluation Metrics:**  
     Define metrics based on the nature of the task (e.g., classification accuracy, F1 score, perplexity). Monitor convergence behavior and computational resource usage.
   - **Cross-Validation and Benchmarking:**  
     Use cross-validation and hold-out test sets to ensure robustness of performance assessments.

5. **Result Analysis and Reporting:**
   - **Comparative Analysis:**  
     Analyze and interpret the performance differences among the RNN, LSTM, GRU, and RNN encoder-decoder models. Assess the trade-offs in terms of processing speed, accuracy, and stability.
   - **Discussion of Findings:**  
     Provide a discussion relative to theoretical expectations (e.g., how and why LSTMs may outperform vanilla RNNs in handling long-term dependencies).
   - **Documentation:**  
     Compile the findings into a detailed technical report, including visualizations of loss curves, performance metrics over training epochs, and error analysis.

6. **Dissemination:**
   - **Presentation and Peer Review:**  
     Present the results at seminars and workshops to receive feedback.
   - **Final Reporting:**  
     Prepare a final research report and, if applicable, a manuscript for publication in a relevant conference or journal.
